Using an Autoencoder Model to Characterize Air Pollution

### 2022-02-08 paper-draft/paper.tex
Fixed autoencoder training function; updated paper

\documentclass{article}
\usepackage[backend=biber, style=authoryear-icomp]{biblatex}
\usepackage{blindtext}
\usepackage{multicol}
\usepackage[document]{ragged2e}
\usepackage[pass, letterpaper]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{subcaption}
\addbibresource{ref.bib}

@@

\begin{abstract}
Climate change is one of the most complex challenges humanity has ever faced. It is a
multi-dimensional problem that poses economic, social, scientific, political, and 
moral questions that must be addressed. One of the most important ways of 
combating climate change is to research pollution mitigation strategies 
so that policy makers and government planners are better equipped to  make 
informed decisions. We focused on the principle cause of global warming:
the polluting gases responsible for the rapid shift in global climate.
We characterize cities by the amount of pollution present using an undercomplete autoencoder
model. By representing the level of eight toxic gases present in numerous cities
across the United States, we explore some of the more prominent features 
that characterize urban pollution. We compare two techniques for reducing 
the dimensionality of our input - Principle Component Analysis (PCA) and
Autoencoders. The first two hidden dimensions represented
a large percentage of the explained variance in the model, so we cluster
the cities in an effort to extract any useful features. 
\end{abstract}
\vspace{0.25cm}
\section{Introduction}

\par Over the past several decades, air pollution has been a major concern for human health
and the environment. The EPA sets national air quality standards for  six major air
toxins: ground level Ozone (O$_3$), Carbon Monoxide (CO), Sulfur Dioxide (SO$_2$),
Lead (Pb), Nitrogen Dioxide (NO$_2$), and particulate matter (PM). The sources of these emissions vary, but the two largest contributors are industrial processes and motorized vehicles. \parencite{NAAQS09}.
\par We wanted to explore how various polluting gases can be characterized with respect to location in an effort to provide researchers and policy makers with the context and tools necessary to make informed decisions. By creating an embedding of time series pollution data for numerous cities throughout the United States, we explore the more interesting features of the model as well as apply it to tangential areas such as the pollution generated by industrial farming and animal husbandry activities.

@@

research. The two major forces behind its competency are, the fact that large volumes of

@@

% Correlation Matrix 
\begin{figure}[h!]
\includegraphics[width=\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/corr_matrix.png}
\caption{Correlation Matrix}
\label{fig:matrix}
\end{figure}

@@

\newpage
\begin{Center}
\begin{figure}[h!]
\caption{Data Pipeline}
\label{fig:pipeline}
    \hspace{1.5cm}\includegraphics[width=10cm, height=13cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pipeline.png}
\end{figure}
\end{Center}

@@

% dataset characteristics table
\begin{table}[h!]
    \caption{Dataset Characteristics}
    \label{tab:table1}
    \vspace{0.1cm}
    \begin{tabular}{p{4cm}p{7cm}}
        \hline
        \multicolumn{2}{c}{OpenWeather Pollution API} \\
        \hline
        Sampling period  & 11/27/20 to 06/05/21\\
        Time interval & Daily  \\
        Cities & 18,526 \\
        Features & 190 \\
        Component gases & 8 \\
        Dependent variable & Last day of the period (06/05/21) \\
        \hline
    \end{tabular}
\end{table}
\par OpenWeather's Pollution API ordinarily returns hourly values, but we opted to use the daily average pollution levels as our feature set in order to make the number of starting dimensons more manageable. The daily averages were also feature engineered to be proportional to the number of features so as to avoid the curse of dimensionality \parencite{Trunk79}. 
\par The original data was unit normalized so that the gradient descents could converge more quickly while still preserving the original differences in range. Each step was saved separately to provide expeditious access to both the normalized and non-normalized data.

@@

\begin{table}[h!]
    \caption{Model Splits}
    \label{tab:table1}
    \vspace{0.1cm}
    \begin{tabular}{p{4cm}p{7cm}}
        \hline
        \multicolumn{2}{c}{Constant Hyperparameters}\\
        \hline
        Training Size & 14,080\\
        Testing Size & 3,705\\
        Validation Size & 741\\
        Total Size & 18,526\\
        \hline
    \end{tabular}
\end{table}
\par A grid search was performed for each model to determine the optimal hyperparameters. The following hyperparameters were tested:
\begin{itemize}
    \item \textit{Learning Rate:} 0.0001, 0.001, 0.01, 0.1
    \item \textit{Batch Size:} 32, 64, 128, 256
    \item \textit{Epochs:} 10, 50, 75, 100
\end{itemize}
\par K-folds cross validation was implemented to validate the grid search using five folds. Additional, the training split for each dimension was separated into a validation set. Each hyperparameter was selected based on the highest coefficient of determination that appeared in each of the five folds for every given dimension. The most frequently occuring combination of hyperparameters was then used to create the finalized embedding. The lowest coefficients of determination were saved and compared with the highest values to give an upper and lower bound of expected values.
\par After some preliminary testing, it was decided that the activation, optimizer, and loss functions should remain constant. The exponential increase in computing time this would have on the grid search drastically outweighed any minor performance gains that a thorough optimization of these functions may have yielded.

\begin{table}[h!]
    \caption{Constant hyperparameters used across all autoencoder models}
    \label{tab:table1}
    \vspace{0.1cm}
    \begin{tabular}{p{4cm}p{7cm}}
        \hline
        \multicolumn{2}{c}{Constant Hyperparameters}\\
        \hline
        Activation function & tanh\\
        Loss function & Mean Squared Error  \\
        Optimizer & Adam \\
        \hline
    \end{tabular}
\end{table}
Moreover, grid search was only performed on a select number of key dimensions rather than on all one hundred ninety dimensions. The first ten dimensions, as well as every tenth dimension until one hundred-twenty were tuned. For each intermediary dimension that did not have a set of key hyperparameters, the hyperparameters of the previous key dimension was used. Once again, the tradeoff between minor performance gains and large compute times led us to adjust the scope of the  hyperparameter tuning.

\begin{table}[h!]
    \caption{Grid Search Results for Carbon Monoxide}
    \label{tab:table2}
    \vspace{0.1cm}
    \begin{tabular}{p{2cm}p{2cm}p{2cm}p{2cm}p{2cm}}
        \hline
        \multicolumn{5}{c}{CO Grid Search Key Dimensions 1-10}\\
        \hline
        dimension & $r^{2}$ & learning rate & batch & epochs  \\
        1 & 0.22 & 0.01 & 64 & 10 \\
        2 & 0.44 & 0.0001 & 256 & 50\\
        3 & 0.49 & 0.001 & 64 & 50\\
        4 & 0.52 & 0.0001 & 32 & 50\\
        5 & 0.56 & 0.0001 & 32 & 100\\
        6 & 0.57 & 0.01 & 32 & 10\\
        7 & 0.60 & 0.0001 & 32 & 150\\
        8 & 0.63 & 0.0001 & 64 & 150\\
        9 & 0.62 & 0.0001 & 32 & 150\\
        10 & 0.62 & 0.0001 & 64 & 100\\
        \hline
    \end{tabular}
\end{table}
\begin{table}[h!]
    \caption{Grid Search Results for Carbon Monoxide}
    \label{tab:table2}
    \vspace{0.1cm}
    \begin{tabular}{p{2cm}p{2cm}p{2cm}p{2cm}p{2cm}}
        \hline
        \multicolumn{5}{c}{CO Grid Search Key Dimensions 20-120}\\
        \hline
        dimension & $r^{2}$ & learning rate & batch & epochs  \\
        20 & 0.62 & 0.01 & 64 & 50\\
        30 & 0.64 & 0.01 & 128 & 75\\
        40 & 0.67 & 0.01 & 256 & 100\\
        50 & 0.68 & 0.01 & 64 & 75\\
        60 & 0.69 & 0.001 & 32 & 150\\
        70 & 0.70 & 0.001 & 64 & 150\\
        80 & 0.71 & 0.001 & 32 & 150\\
        90 & 0.71 & 0.01 & 64 & 100\\
        100 & 0.72 & 0.01 & 64 & 150\\
        110 & 0.72 & 0.01 & 32 & 75\\
        120 & 0.73 & 0.001 & 64 & 100\\
        \hline
    \end{tabular}
\end{table}
\newpage

@@

\section{Results}
We compared the percent explained variance across one hundred ninety dimensions for the encoded values generated by the PCA and autoencoder models. While PCA was overall less noisy, we found that the autoencoder model achieved a much higher explainable variance for the first several dimensions.
\comment{
\\ \note{Note that results graphs are placeholders}
% ae vs pca linegraph

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
    \caption{PCA Dimensionality Reduction}
    \label{fig:ae_dim_reduction}
\end{subfigure}% <- % sign must be here!
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/ae_all_dims.png}
    \caption{PCA Dimensionality Reduction}
    \label{fig:pca_dim_reduction}
\end{subfigure}
\caption{Graph of autoencoder and PCA reduced representations of air pollutants}
\label{fig:ae_vs_pca}
\end{figure}
}

\begin{figure}[h!]
    \includegraphics[width=\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
    \caption{PCA Dimensionality Reduction}
    \label{fig:ae_dim_reduction}
\end{figure}

\begin{figure}[h!]
    \includegraphics[width=\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
    \caption{Autoencoder Dimensionality Reduction\ \note{Placeholder for AE}}

    \label{fig:ae_dim_reduction}
\end{figure}

\par We clustered the cities based on the first two latent dimensions in order to see how
the model characterizes each city

% co scatter
\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/co_scatter_ae_wlist.png} 
    \caption{Outlier Cities}
    \label{fig:outliers}
\end{subfigure}% <- % sign must be here!
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/co_scatter_ae_large_cities.png}
    \caption{Densely Populated Cities}
    \label{fig:dense_cities}
\end{subfigure}

\caption{Scatter plots of outlier and highly-populated cities over first two dimensions}
\label{fig:outliers_vs_dense_cities}
\end{figure}

\par By selecting from a list the top two hundred most highly-populated cities in the US, we found that nearly all of them were located near the center of the clustering. Conversely, we found that many of the outliers were smaller cities in rural locations.

@@

Although the model provides insights on the most succinct features that characterize urban air pollution, more specific external data is needed to uncover these features. One study \parencite{Bai2019} uses a stacked autoencoder approach and focuses on the seasonality of pm2.5 levels. The proposed model considers both the seasonality and temporality of pm2.5 levels and supporting meteorological data. The research group was able to accurately predict the pm2.5 levels to the hour for a given season.

>>>>

\documentclass{article}
\usepackage[backend=biber, style=authoryear-icomp]{biblatex}
\usepackage{blindtext}
\usepackage{multicol}
\usepackage[document]{ragged2e}
%\usepackage[pass, letterpaper]{geometry}
\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{subcaption}
\addbibresource{ref.bib}

@@

\begin{abstract}
Climate change is one of the most complex challenges humanity has ever faced. It is a
multi-dimensional problem that poses economic, social, scientific, political, and 
moral questions that must be addressed. One of the most important ways of 
combating climate change is to research pollution mitigation strategies 
so that policy makers and government planners are better equipped to  make 
informed decisions. We focused on the principle cause of global warming:
the polluting gases responsible for the rapid shift in global climate.
We characterize cities by the amount of pollution present using an undercomplete autoencoder
model. By representing the level of eight toxic gases present in numerous cities
across the United States, we explore some of the more prominent features 
that characterize urban pollution. We compare two techniques for reducing 
the dimensionality of our input - Principle Component Analysis (PCA) and
Autoencoders. The first two hidden dimensions represented
a large percentage of the explained variance in the model, so we cluster
the cities in an effort to extract any useful features. 
\end{abstract}

\vspace{0.25cm}
\section{Introduction}

\par For the past several decades, air pollution has become a major concern for human health
and the environment. The EPA sets national air quality standards for  six major air
toxins: ground level Ozone (O$_3$), Carbon Monoxide (CO), Sulfur Dioxide (SO$_2$),
Lead (Pb), Nitrogen Dioxide (NO$_2$), and particulate matter (PM). The sources of these emissions vary, but the two largest contributors are industrial processes and motorized vehicles. \parencite{NAAQS09}.
\par We wanted to explore how various polluting gases can be characterized with respect to location in an effort to provide researchers and policy makers with the context and tools necessary to make informed decisions. By creating an embedding of time series pollution data for numerous cities throughout the United States, we explore the more interesting features of the model as well as apply it to tangential areas such as the pollution generated by industrial farming and animal husbandry activities.

@@

research. The two major forces behind its competency are the fact that large volumes of

@@

% Correlation Matrix 
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.7\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/corr_matrix.png}
\end{center}
    \caption{Correlation Matrix for Carbon Monoxide (daily averages for cities throughout the US)}
\label{fig:matrix}
\end{figure}

@@

\newpage
\begin{Center}
\begin{figure}[h!]
\caption{Data Pipeline}
\label{fig:pipeline}
    \begin{center}
    \includegraphics[width=10cm, height=13cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pipeline.png}
    \end{center}
\end{figure}
\end{Center}

@@

% dataset characteristics table
\begin{table}[h!]
\begin{center}
    \caption{Dataset Characteristics}
    \label{tab:table1}
    \vspace{0.1cm}
    \begin{tabular}{p{4cm}p{7cm}}
        \hline
        \multicolumn{2}{c}{OpenWeather Pollution API} \\
        \hline
        Sampling period  & 11/27/20 to 06/05/21\\
        Time interval & Daily  \\
        Cities & 18,526 \\
        Features & 190 \\
        Component gases & 8 \\
        Dependent variable & Last day of the period (06/05/21) \\
        \hline
    \end{tabular}
\end{center}
\end{table}
\par OpenWeather's Pollution API ordinarily returns hourly values, but we opted to use the daily average pollution levels as our feature set in order to make the number of starting dimensons more manageable. The daily averages were also feature engineered to be proportional to the number of features so as to avoid the curse of dimensionality \parencite{Trunk79}. 
\par The original data was unit normalized so that the gradient descents could converge more quickly while still preserving the original differences in range. Each step was saved separately to provide expeditious access to both the normalized and non-normalized data.

@@

\begin{table}[h!]
\begin{center}
    \caption{Model Splits}
    \label{tab:table1}
    \vspace{0.1cm}
    \begin{tabular}{p{4cm}p{7cm}}
        \hline
        \multicolumn{2}{c}{Constant Hyperparameters}\\
        \hline
        Training Size & 14,080\\
        Testing Size & 3,705\\
        Validation Size & 741\\
        Total Size & 18,526\\
        \hline
    \end{tabular}
\end{center}
\end{table}
\par A grid search was performed for each model to determine the optimal hyperparameters. The following hyperparameters were tested:
\begin{itemize}
    \item \textit{Learning Rate:} 0.0001, 0.001, 0.01, 0.1
    \item \textit{Batch Size:} 32, 64, 128, 256
    \item \textit{Epochs:} 10, 50, 75, 100
\end{itemize}
\par K-folds cross validation was implemented to validate the grid search using five folds. Additional, the training split for each dimension was separated into a validation set. Each hyperparameter was selected based on the highest coefficient of determination that appeared in each of the five folds for every given dimension. The most frequently occuring combination of hyperparameters was then used to create the finalized embedding. The lowest coefficients of determination were saved and compared with the highest values to give an upper and lower bound of expected values.
\par After some preliminary testing, it was decided that the activation, optimizer, and loss functions should remain constant. The exponential increase in computing time this would have on the grid search drastically outweighed any minor performance gains that a thorough optimization of these functions may have yielded.

\begin{table}[h!]
\begin{center}
    \caption{Constant hyperparameters used across all autoencoder models}
    \label{tab:table1}
    \vspace{0.1cm}
    \begin{tabular}{p{4cm}p{7cm}}
        \hline
        \multicolumn{2}{c}{Constant Hyperparameters}\\
        \hline
        Activation function & tanh\\
        Loss function & Mean Squared Error  \\
        Optimizer & Adam \\
        \hline
    \end{tabular}
\end{center}
\end{table}
Moreover, grid search was only performed on a select number of key dimensions rather than on all one hundred ninety dimensions. The first ten dimensions, as well as every tenth dimension until one hundred-twenty were tuned. For each intermediary dimension that did not have a set of key hyperparameters, the hyperparameters of the previous key dimension was used. Once again, the tradeoff between minor performance gains and large compute times led us to adjust the scope of the  hyperparameter tuning.

\begin{table}[h!]
\begin{center}
    \caption{Grid Search Results for Carbon Monoxide}
    \label{tab:table2}
    \vspace{0.1cm}
    \begin{tabular}{p{2cm}p{2cm}p{2cm}p{2cm}p{2cm}}
        \hline
        \multicolumn{5}{c}{CO Grid Search Key Dimensions 1-10}\\
        \hline
        dimension & $r^{2}$ & learning rate & batch & epochs  \\
        1 & 0.22 & 0.01 & 64 & 10 \\
        2 & 0.44 & 0.0001 & 256 & 50\\
        3 & 0.49 & 0.001 & 64 & 50\\
        4 & 0.52 & 0.0001 & 32 & 50\\
        5 & 0.56 & 0.0001 & 32 & 100\\
        6 & 0.57 & 0.01 & 32 & 10\\
        7 & 0.60 & 0.0001 & 32 & 150\\
        8 & 0.63 & 0.0001 & 64 & 150\\
        9 & 0.62 & 0.0001 & 32 & 150\\
        10 & 0.62 & 0.0001 & 64 & 100\\
        \hline
    \end{tabular}
\end{center}
\end{table}
\begin{table}[h!]
\begin{center}
    \caption{Grid Search Results for Carbon Monoxide}
    \label{tab:table2}
    \vspace{0.1cm}
    \begin{tabular}{p{2cm}p{2cm}p{2cm}p{2cm}p{2cm}}
        \hline
        \multicolumn{5}{c}{CO Grid Search Key Dimensions 20-120}\\
        \hline
        dimension & $r^{2}$ & learning rate & batch & epochs  \\
        20 & 0.62 & 0.01 & 64 & 50\\
        30 & 0.64 & 0.01 & 128 & 75\\
        40 & 0.67 & 0.01 & 256 & 100\\
        50 & 0.68 & 0.01 & 64 & 75\\
        60 & 0.69 & 0.001 & 32 & 150\\
        70 & 0.70 & 0.001 & 64 & 150\\
        80 & 0.71 & 0.001 & 32 & 150\\
        90 & 0.71 & 0.01 & 64 & 100\\
        100 & 0.72 & 0.01 & 64 & 150\\
        110 & 0.72 & 0.01 & 32 & 75\\
        120 & 0.73 & 0.001 & 64 & 100\\
        \hline
    \end{tabular}
\end{center}
\end{table}
\newpage

@@

\newpage

\section{Results}

\par We compared the percent explained variance across all one hundred ninety dimensions for the encoded values generated by the PCA and AE models. While PCA was overall less noisy, we found that the autoencoder model achieved a much higher explainable variance for the first several dimensions.

\begin{center}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.7\linewidth, height=7cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
    \caption{PCA Dimensionality Reduction}
    \label{fig:pca_dim_reduction}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.7\linewidth, height=7cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
    \caption{Autoencoder Dimensionality Reduction\ \note{Placeholder for AE}}
    \label{fig:ae_dim_reduction}
\end{figure}
\end{center}

\par In nearly every case shown in figures \ref{fig:pca_dim_reduction} and \ref{fig:ae_dim_reduction}, the autoencoder model shows consistently higher explainability than the PCA. Tropospheric Ozone (O3) in particular shows a massive jump of over fifty percentage points explainable variance in the first two dimensions. For this reason, O3 was the primary model for the subsequent visualizations.
\par In order to better visualize how the hidden layers distill information, we clustered the cities based on the first two encoded dimensions. Additionally, a number of location heuristics were applied to further understand the distribution of cities. Figure \ref{fig:heuristic_labels} shows how the embeddings are clustered by regions, as well as by individual states and territories.

% co scatter
\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/region.png} 
    \caption{Model characterization grouped by region}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/outside-us.png}
    \caption{Model characterization of U.S. Territories}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/california.png} 
    \caption{Model characterization of California}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/texas.png} 
    \caption{Model characterization of Texas}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/new-york.png}
    \caption{Model characterization of New York}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/minnesota.png}
    \caption{Model characterization of Minnesota}
\end{subfigure}
\caption{Scatter plots of outlier and highly-populated cities over first two dimensions}
\label{fig:heuristic_labels}
\end{figure}

\par The model's characterization appears to be heavily dependent on geographic location. When grouped by region, one can see that there is a fairly tight distribution, especially among the Midwest and Northeast regions. Puerto Rico is also tightly clustered, but the territory contains a limited 211 representative data points.

\par Curiously enough, population size does not seem to be one of the most succint features of the model. Figure \ref{fig:population_comparison} shows the top two hundred most populous cities in the US for the first four dimensions of the model. The first four latent dimensions of O3 account for nearly 60\% of the total explainable variance. However, the top 200 cities appear to be randomly dispersed throughout, suggesting that population alone has less of an effect on the model. It is possible that population is represented indirectly

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.0\linewidth, height=10cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/population.png} 
    \caption{Model characterization grouped by region}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.0\linewidth, height=10cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/population_3and4.png}
    \caption{Model characterization of U.S. Territories}
\end{subfigure}
\caption{Scatter plots of outlier and highly-populated cities over first two dimensions}
\label{fig:population_comparison}
\end{figure}

\par It is difficult to determine why the model does not seem to consider population size. One reason for this may be that temperature and climate "overpowers" any influence a city's size may have. Any number of secondary effects related to climate and location could pull the model away from considering population. Moreover, The time frame of our data spans from winter 2020 to early summer of 2021, which is not long enough to fully capture seasonality. Given a larger data set, it is likely that both population and seasonality would play much larger roles.

\par Due to the importance of geography in the embedding, we created several geographic scatter plots to visualize both the embedding values and pollution levels with respect to location.

% Scattergeo 
\begin{figure}[h!]
    \caption{Geographic comparison of model embedding values versus daily pollution levels}
    \label{fig:outliers_vs_dense_cities}
\begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-dim1.png} 
    \caption{O3 values of the first encoded dimension (units of compression)}
    \label{fig:outliers}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-first.png}
    \caption{O3 pollution values for the first day of data set ($\mu g/m^3$)}
    \label{fig:dense_cities}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-dim2.png} 
    \caption{O3 values of the second encoded dimension (units of compression)}
    \label{fig:outliers}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-last.png}
    \caption{O3 pollution values for the last day of data set ($\mu g/m^3$)}
    \label{fig:dense_cities}
\end{subfigure}
\end{figure}

\newpage

@@

\par The proposed model provides us with many insights on the most succinct features that characterize urban air pollution in US cities. Based on location identifiers such as city name, state, country, and the geograpic coordinates, we are able to infer much of the explainability of the model. However, more specific external data is needed to fully uncover these features. One study \parencite{Bai2019} uses a stacked autoencoder approach and focuses on the seasonality of pm2.5 levels. The model considers both the seasonality and temporality of pm2.5 levels and supporting meteorological data. The research group was able to accurately predict the pm2.5 levels to the hour for a given season. The pollution autoencoder models we created only consider half a year of data, so we were unable to fully account for seasonal variations in the dataset.
\par Another study focusing on pm2.5 \parencite{doreswamy2021} uses autoencoders for spatio-temporal clustering analysis. The group looks at Taiwan for its massive air pollution problem, coupled with the fact that the area's unique topography makes it difficult to map the distribution of pollution. The model uses aproximately six years of censor data from Eastern Taiwan to analyze the spatio-temporal qualities of pm2.5, as well as the socioeconomic effects of its presence.
